MADA Data Analysis Project

### 2021-11-13 code/analysis_code/03-P2.Data.rmd
Update References
Select and add References to manuscript

```{r}
#save data frame table to file for later use in manuscript
summarytable_file <- here::here("products","Project_Parts","Part_3", "P2.Summary.Table.rds")
saveRDS(P2summary_df, file = summarytable_file)
```

>>>>

```{r}
#save data frame table to file for later use in manuscript
summarytable_file <- here::here("products","Project_Parts","Part_3", "P2.Summary.Table.rds")
#saveRDS(P2summary_df, file = summarytable_file)
```

### 2021-11-13 products/manuscript/Manuscript.Rmd
Update References
Select and add References to manuscript

---
title: "MADA Data Analysis Project"
subtitle: ""
author: Monica Chan
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
documentclass: article
output:
  bookdown::word_document2: 
    toc: false
    number_sections: true
  bookdown::html_document2: 
    toc: false
bibliography: ../references.bib
csl: ../apa.csl
---

# Summary/Abstract
_Write a summary of your project._  
Applying MADA learned techniques to a RT-PCR dataset to determine the effectiveness of sample enhancement strategies for environmental surface sampling of low bio-burden viruses.

@@

##  Background
The global Coronavirus Disease 2019 (COVID-19) pandemic continues to persist due to the timing of an available vaccine, varying adherence to COVID-19 public health safety recommendations, and the emergence of SARS-CoV-2 (SC2) variants. Environmental reservoirs are a reported, indirect transmission route for viral pathogens. Although many viruses have a low infectious dose, they can be difficult to detect due to the low bioburden in the environment. The ability to detect viral pathogens is critical to identifying hidden reservoirs which can inform infection control. Culture-based methods can be labor and time intensive. The use and limitations of commercial “target capture” (TC) technologies are varied compared to molecular approaches utilizing automated extraction. The objective of this study is improve outbreak response through rapid pathogen detection by evaluating a TC technology’s ability to capture SC2 from complex matrices.

@@

Using established standard methods for environmental sampling with a self-contained polyurethane swab from stainless steel surface, a TC approach (NTP; Nanotrap® Magnetic Virus Particles, Ceres Nanosciences) was compared to an automated extraction method (MX48; Maxwell® RSC Viral Total Nucleic Acid Purification Kit, Promega) for recovery of heat-inactivated SC2 from the polyurethane swab as detected by rRT-PCR

@@

processingscript.R was modified to do initial cleaning and preparation of the dataset for further exploration.

@@

_Use a combination of text/tables/figures to explore and describe your data. You should produce plots or tables or other summary quantities for the most interesting/important quantities in your data. Depending on the total number of variables in your dataset, explore all or some of the others.

_Continue by creating plots or tables of the outcome(s) of interest and the predictor/exposure/input variables you are most interested in. If your dataset is small, you can do that for all variables. Plots produced here can be scatterplots, boxplots, violinplots, etc. Tables can be simple 2x2 tables or larger ones._

_To get some further insight into your data, if reasonable you could compute simple statistics (e.g. t-tests, simple regression model with 1 predictor, etc.) to look for associations between your outcome(s) and each individual predictor variable. Though note that unless you pre-specified the outcome and main exposure, any "p<0.05 means statistical significance" interpretation is not valid._

@@

Summary table 
scatter plot of SDs based ond dilution factors

#### Phase 1

@@

@@

_Use one or several suitable statistical/machine learning methods to analyze your data and to produce meaningful figures, tables, etc. This might again be code that is best placed in one or several separate R scripts that need to be well documented. You want the code to produce figures and data ready for display as tables, and save those. Then you load them here._

@@

## Summary and Interpretation
_Summarize what you did, what you found and what it means._

## Strengths and Limitations
_Discuss what you perceive as strengths and limitations of your analysis._

## Conclusions
_What are the main take-home messages?_

_Include citations in your Rmd file using bibtex, the list of references will automatically be placed at the end_

This paper [@Leek2015a] discusses types of analyses.

Note that this cited reference will show up at the end of the document, the reference formatting is determined by the CSL file specified in the YAML header. Many more style files for almost any journal [are available](https://www.zotero.org/styles). You also specify the location of your bibtex reference file in the YAML. You can call your reference file anything you like, I just used the generic word `references.bib` but giving it a more descriptive name is probably better.

>>>>

---
title: "MADA Data Analysis Project"
subtitle: ""
author: Monica Chan
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
documentclass: article
output:
  bookdown::html_document2: 
    toc: false
  bookdown::word_document2: 
    toc: false
    number_sections: true
bibliography: ../references.bib
csl: ../apa.csl
---

# Summary/Abstract

Applying MADA learned techniques to a RT-PCR dataset to determine the effectiveness of sample enhancement strategies for environmental surface sampling of low bio-burden viruses.

@@

##  Background
The global Coronavirus Disease 2019 (COVID-19) pandemic continues to persist due to the timing of an available vaccine, varying adherence to COVID-19 public health safety recommendations, and the emergence of SARS-CoV-2 (SC2) variants. Environmental reservoirs are a reported, indirect transmission route for viral pathogens [@Marcenac2021-sz]. Although many viruses have a low infectious dose, they can be difficult to detect due to the low bioburden in the environment. The ability to detect viral pathogens is critical to identifying hidden reservoirs which can inform infection control [@Paton2021-jc]. Culture-based methods can be labor and time intensive. The use and limitations of commercial “target capture” (TC) technologies are varied compared to molecular approaches utilizing automated extraction. The objective of this study is improve outbreak response through rapid pathogen detection by evaluating a TC technology’s ability to capture SC2 from complex matrices.

@@

Using established standard methods for environmental sampling with a self-contained polyurethane swab from stainless steel surface, a TC approach (NTP; Nanotrap® Magnetic Virus Particles, Ceres Nanosciences) was compared to an automated extraction method (MX48; Maxwell® RSC Viral Total Nucleic Acid Purification Kit, Promega) for recovery of heat-inactivated SC2 from the polyurethane swab as detected by rRT-PCR [@Wittwer2012-vt][@Shah2021-mb].

@@

processing script.R was modified to do initial cleaning and preparation of the data set for further exploration.

@@

Direction was influenced by [@Weir2016-sj] on "Effect fo Surface Sampling and Recovery of Viruses and No-Spore-Forming Bacteria on a Quantitative Microbial Risk Assessment Model for Fomites".

Atempt to do as [@Brandolini2021-ja]

@@

Summary table 
scatter plot of SDs based on dilution factors
```{r}
knitr::include_graphics("../../results/ALL-SD-Plot.png")
```
#### Phase 1

```{r}
knitr::include_graphics("../../results/P1-SD-Plot.png")
```

@@

```{r}
knitr::include_graphics("../../results/P2-SD-Plot.png")
```
#### Phase 2 Comparing A and B Tool Recovery 
```{r}
knitr::include_graphics("../../results/ASK-P2P1-Tools.png")
```
Tool A has an average Ct recovery higher than B. This makes sense as the surfaces A recovered from is 2 logs higher (-1) than that of B (-3). A log drop is ~3Ct values we can see from the graphic that it trends as predicted. Suggests that

#### Phase 2 Comparing Method Serial Dilutions (CDC vs NTP) with standard input of 300uL
```{r}
knitr::include_graphics("../../results/ASK-P2P2-Methods.png")
```
Serial dilution data between CDC and NTP. Noting that all SD are in the confirmed realm of under 27 Ct Value CDC method is better than NTP at -1 concentration of the SD. the -2 shows little difference between method recovery Ct values. NTP appears to remain better than CDC at -3 and -4 levels. Accounting for the SE looks like both methods are the same until the -2 dilution. Might be significant difference at the -4 dilution.

#### Phase 2 Comparing Methods with standard input of 300uL
```{r}
knitr::include_graphics("../../results/ASK-P2P3-Methods0300.png")
```

#### Phase 2 Method extraction comparisons with all inputs and tools
```{r}
knitr::include_graphics("../../results/ASK-P2P4-MethodsALL.png")
```

#### Phase 2 Nanoparticles only - Compare detectors with dilution and input in Serial Dilutions
```{r}
knitr::include_graphics("../../results/ASK-P2P5-NTP-SD.png")
```

#### Phase 2 Nanoparticles only - All tools at 300 and 1000ul input levels
```{r}
knitr::include_graphics("../../results/ASK-P2P6-NTP.png")
```

@@

@@

## Summary and Interpretation
Took

## Strengths and Limitations
Need more replicates, but this is a time consuming processes just to get the data.

Additional materials need to be considered [@Jansson2020-cg].

## Conclusions
